Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2015-07-05 Paper/paper.tex
updates to the paper

\item For each system comprised of all unique combinations of the above 
parameter choices, we carry out two parameter estimation tests: 
\begin{enumerate}
\item NS-BH signals injected, recovered with BH-BH waveforms, and
\item NS-BH signals injected, recovered with NS-BH waveforms.
\end{enumerate}
Injections were made at different SNR values: 
\begin{enumerate}
\item $\rho = \{20, 30, 40, 50, 60, 70, 80, 90, 100\}$.
\end{enumerate}
Choice of emcee parameters:
\begin{enumerate}
\item $N_\mathrm{samples}=6000$; 
\item $N_\mathrm{walkers}=100$; 
\item $N_\mathrm{burn-in}=500$; 
\end{enumerate}
and choice of prior boundaries:
\begin{enumerate} 
\item $m_\mathrm{BH}\in [1.2, 25]M_\odot$;
\item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
\item $\Lambda\leq 2000$; 
\item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
\end{enumerate}

@@

\item We deduce from
\begin{enumerate}
\item large chirp mass biases, 
\item lack of convergence of $90\%$ confidence intervals with $\rho$ (except for
$\Lambda$, noticeably), and
\item jagged-ness of parameter bias curves against $\rho$
\end{enumerate}
that the MCMC chains have not yet converged.
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modeling tidal deformation during inspiral}
Describe the PN terms used here

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SNR loss from ignoring NS tidal effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
See what is the loss in SNR if we recover NSBH signals with a template bank of
BH-BH waveforms? It is very likely that the answer would be: $<1\%$. But it
is very interesting to show that for spinning BHs!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Systematic bias in recovered parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[h]

@@

\textbf{With BH-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
\textbf{With NS-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
\caption{These figures show the (fractional) difference between the median value of
recovered posterior distribution for binary chirp mass and the actual injected chirp mass,
as a function of the signal-to-noise-ratio (SNR) of the injected signal.
The top and bottom panels show results for parameter estimation performed using
waveform templates without and with NS matter effects included, respectively.
In each row, the left and right panel correspond to 
$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
In each panel, the color of the curves correspond to the value of NS deformability 
parameter, and line style correspond to the dimensionless spin on the BH 
($\chi_\mathrm{BH}\equiv \dfrac{a}{m_\mathrm{BH}^2}$).
}
\label{fig:TNT_chirpMassBias_vs_SNR_q23}
\begin{figure*}[h]
	\textbf{With BH-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
	\textbf{With NS-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
	\caption{These figures show the narrowing of the $90\%$ confidence 
		interval (recovered) for chirp mass, as a function of the 
		signal-to-noise-ratio (SNR). Note that the confidence interval
                width is normalized by the true parameter value.
                The binary parameters, i.e. black hole spin,
                mass-ratio, NS tidal deformability, are all depicted in the 
                same way as in Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}.
	}
	\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
\begin{figure*}[h]
\textbf{With BH-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
\textbf{With NS-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}%\quad
\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},

@@

mass-ratio $\eta$ here.}

@@

\begin{figure*}[h]

@@

	\textbf{With BH-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
	\textbf{With NS-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}%\quad
	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},

@@

		of the dimensionless
		mass-ratio $\eta$ here.}

@@

\textbf{With BH-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
\textbf{With NS-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}%\quad
\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},

@@

fractions of the injected spin.}

@@

	\textbf{With BH-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
	\textbf{With NS-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}%\quad
	\caption{This figure is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},

@@

		of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.}

@@

\textbf{With NS-BH Templates}
\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}

@@

	\textbf{With NS-BH Templates}
	\includegraphics[width=2\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}

@@

include the effects of the NS tidal deformability. While this is not expected to

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using templates with tidal effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having shown in the previous section that we begin to care for NS tidal effects
for signals in the XXX corner of the paraemter space, with SNRs above YY, here
we investigate the effect of the improvement in the accuracy of the recovered
parameters when tidal effects are included in the templates.

\begin{enumerate}
\item What is the reduction in the bias of the maximum likelihood parameters
when using tidal templates?\newline
``Plot the ratio of the bias between N templates and T templates (against T
signals), as a function of SNR.''
\item What is the reduction in the uncertainty in binary mass and spin, if
any?\newline
``show how the the $95\%$ confidence intervals shrink, as a function of SNR, 
when we go from using N templates to T templates.''
\end{enumerate}

>>>>

\item For each system comprised of all unique combinations of the above 
parameter choices, we carry out three parameter estimation tests: 
\begin{enumerate}
\item BH-BH signals injected, recovered with BH-BH waveforms (control),
\item NS-BH signals injected, recovered with BH-BH waveforms, (current approach) and
\item NS-BH signals injected, recovered with NS-BH waveforms.
\end{enumerate}
Injections were made at different SNR values: 
\begin{enumerate}
\item $\rho = \{20, 30, 50, 70, 90, 120\}$.
\end{enumerate}
Choice of emcee parameters:
\begin{enumerate}
\item $N_\mathrm{samples}=150,000$; 
\item $N_\mathrm{walkers}=100$; 
\item $N_\mathrm{burn-in}=500$; 
\end{enumerate}
and choice of prior boundaries:
\begin{enumerate} 
\item $m_\mathrm{BH}\in [1.2, 25]M_\odot$;
\item $m_\mathrm{NS}\in [1.2, 15]M_\odot$;
\item $\Lambda\leq 4000$; 
\item $\sigma\left(\Lambda\right) = 100$ (for chains where templates are NS-BH waveforms).
\end{enumerate}

@@

\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Modeling tidal deformation during inspiral}
Describe the PN terms used here

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How are detection searches affected by not including NS matter effects}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
See what is the loss in SNR if we recover NSBH signals with a template bank of
BH-BH waveforms? It is very likely that the answer would be: $<1\%$. But it
is very interesting to show that for spinning BHs!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How is PE affected by not including NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering 
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassBias_vs_SNR_q23.pdf}%\quad
\caption{These figures show the (fractional) difference between the chirp mass 
value corresponding to the median of its posterior probability distribution,
and the actual injected chirp mass,
as a function of the signal-to-noise-ratio (SNR) of the injected signal.
% In the top row, NS matter effects on the inspiral and merger waveform
% are neithed included in templates nor in the injected signal. 
In the top row, NS matter effects (inspiral and merger) are included in the
injections but not in templates. This corresponds to the current 
plan for aLIGO parameter estimation studies. The bottom row shows the effect
of additionally including NS matter effects in the templates.
% 
In each row, the left and right panel correspond to 
$q=m_\mathrm{BH}/m_\mathrm{NS}=\{2,3\}$, respectively. 
In each panel, the color of each curve corresponds to the value of NS 
deformability parameter $(\Lambda=\{500,1000,2000\})$, 
and line-style corresponds to the value of
the dimensionless spin on the BH ($\chi_\mathrm{BH}=\{-0.5,0,+0.5\}$).
}
\label{fig:TNT_chirpMassBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering 
% \textbf{BH-BH Injection; BH-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/NN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_chirpMassCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_chirpMassCIWidth90_vs_SNR_q23.pdf}%\quad
\caption{These figures are similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}
with the difference that here we show the width of the $90\%$ confidence 
interval (recovered) for chirp mass, as a function of the injected SNR. 
Note that the confidence interval's width is normalized by the injected
parameter value.
}
\label{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\newpage
\newpage

\begin{figure*}[!t]

@@

\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},

@@

mass-ratio $\eta$ here. In the bottom row we show control runs where both
signal and template waveforms lack matter effects.}

@@

@@

\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},

@@

  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
  confidence interval is shown here.}

@@

\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
\caption{The top two rows in this figure are
similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},

@@

fractions of the injected spin. In the bottom row, we show a control run where
both injected and template waveforms correspond to binary black holes.}

@@

\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
\caption{The top two rows in the figure are similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},

@@

of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
width of the confidence interval without normalizing with the injected spin.
The bottom row show a control run where matter effects are ignored both 
in the injected signal and the template waveforms.}

@@

\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}

@@

\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}

@@

include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What do we gain by using templates that include NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having shown in the previous section that we begin to care for NS tidal effects
for signals in the XXX corner of the paraemter space, with SNRs above YY, here
we investigate the effect of the improvement in the accuracy of the recovered
parameters when tidal effects are included in the templates.

\begin{enumerate}
\item What is the reduction in the bias of the maximum likelihood parameters
when using tidal templates?\newline
``Plot the ratio of the bias between N templates and T templates (against T
signals), as a function of SNR.''
\item What is the reduction in the uncertainty in binary mass and spin, if
any?\newline
``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
when we go from using N templates to T templates.''
\end{enumerate}

\red{Could have subsections for aLIGO and ET ?}
